\chap{Acknowledgements}

%%%% there is no specific formatting requirement for this section 
%% default is double-spaced (you can style it as you want)
While this dissertation has myself listed as the sole author on the title page, the research detailed within is the product of many individuals. Foremost, I am deeply indebted to the my advisor, mentor, and friend, Paul Smolensky. When I initially reached out about pursuing a Master's degree under his supervision, Paul indicated that he had already accepted his final student in preparation of transitioning to an emeritus role. However, after a brief correspondence about my research interests, he decided to make an exception—an exception that, unbeknownst to me at the time, he had made twice prior. Paul's mentorship and guidance have shaped my research profoundly, and I am immensely grateful that he extended his retirement to see me through my doctorate.

In addition to making an exception to take just one more student, I am thankful that Paul also made the decision to accept two other ``final'' students: Tom McCoy and Najoung Kim. Tom and Najoung were instrumental in helping me adjust to graduate school, teaching me how to conduct research, and inspiring me to think like a computational linguist. I am also grateful for the stimulating conversations and discussions with members of Paul's Neurosymbolic Computation Lab: Caitlin Smith, Eric Rosen, Gabriel Kressin-Palacios, Matthias Lalisse, and Coleman Haley. Additionally, Leyla Isik provided invaluable mentorship during my second project, complemented by the friendship and community of her Computational Cognitive Neuroscience Lab.

I was fortunate to complete internships at Microsoft Research and IBM Research, both of which significantly enriched my doctoral work and better prepared me for a career in industry. My collaboration with the Microsoft Research Deep Learning group extended beyond my internships, and I am especially grateful to Roland Fernandez and Jianfeng Gao for making sure that I had the resources necessary to conduct the experiments described in this dissertation. At IBM, I deeply appreciate Abbas Rahimi’s mentorship and the fun and diverse environment of the In-Memory Computing Group, especially Aleksandar Terzić, Michael Hersche, and Abu Sebastian. In particular, team lunch was a highlight of each work day, especially the opportunity to discuss topics and hear viewpoints from so many different countries.

My doctoral research benefited greatly from collaboration with an outstanding group of co-authors: Tal Linzen, Yichen Jiang, Asli Celikyilmaz, Sudha Rao, Hamid Palangi, Mohit Bansal, Edward Hu, Kate McCurdy, Yunmo Chen, Henry Conklin, and Mattia Opper. I also sincerely appreciate the insightful feedback on preprints provided by Michael Bonner, Emalie McMahon, Raj Magesh, Ricky Loynd, and Steven Piantadosi.

This dissertation is anchored by two bookends. Initially, I was guided into research by Tom Griffiths, Joshua Peterson, and Aida Nematzadeh, who generously mentored me when I was an eager software engineer with no research experience. Tom, in particular, indirectly set me on my doctoral path through a parenthetical citation to Paul’s work in \citet{griffiths_probabilistic_2010}. At the other end, I am tremendously grateful to the members of my dissertation committee for the time they spent improving my work: Benjamin Van Durme, John T. Hale, Colin Wilson, Robert Frank, Jennifer Hu, and Christopher Honey.

My experience at Hopkins was profoundly enriched by the vibrant intellectual communities within the Cognitive Science, Psychology, and Computer Science departments, as well as the Center for Speech and Language Processing. I am grateful to Géraldine Legendre for her leadership as Chair of the Cognitive Science Department, and to Sarah Ciotola, Chamera Sampson, Sue Potterfield, and Peggy MacKenzie, whose tireless administrative efforts ensured the timely completion of this dissertation.

Friendship has provided crucial balance during my doctoral journey. I am incredibly fortunate to remain friends with so many people from High School, and our Collegiate family has only grown larger since then with the addition of partners, children, and pets, as well as my enduring friendships with Gabbie and Sarah. My undergraduate friendships with Charlie, Nicky, and Andrew have become even more profound with time, while my first stint in California was enriched by friendships with Drake, Brittany, Gatsby, Kira, Robin, Gable, Johnny, Wyatt, Will, Emily, and Jed. In Baltimore, I cherised the time that I got to spend with Matt, Lauren, and Mousse, and the unforgettable family dinners hosted with David, Mimi, Chris, and Karli, rank among my fondest memories. Special thanks to Sable, Bran, Saoirse, and Thea for being trustworthy neighbors; to Kelly and Andrea for teaching me to savor joy; and to Daniel and Hannah for sharing art and music.

Finally, I owe an immeasurable debt of gratitude to my family for their unwavering support and the huge role that they played in my development and journey. Thank you to Mom and Dad for all of the support you have provided throughout my life—in retrospect, your combined interests in electrical engineering and neurology shaped my intellectual curiosities from the very start. Pam, your constant presence has always been a source of comfort, and welcoming Holly into our family marked one of the happiest moments of my life. Expanding outwards, I am thank for the support that my grandparents provided, and to my cousins, aunts, uncles, and other close relatives for the delightful times at our family gatherings. Irina (and Pepper, obviously), your love and companionship have been my greatest joy and motivation of each day, and you have been on my mind for every second of the past four years. I am so excited to build our life together.